\large{\bf Looking forwards and backwards: dynamics and genealogies of locally regulated populations }

### 2023-12-08 review-responses.tex

\begin{point}{\revref}
The first paragraph seems sloppy. For example, this expression [for $\mu_\theta(x)$] is non-negative for
what range of values for $(x, \eta)$? Why ``anyway true for sufficiently large $\theta$''? Perhaps this
assumption should be labeled and called upon whenever is it being referred to.
\end{point}

\reply{
}

\begin{point}{\revref}
Definition \ref{def:lineage_generator}, Proposition 2.23: I think these may need to be rewritten/rephrased. Up
to Proposition \ref{thm:lineages}, it is not clear what ``that individual'' is in the limiting model. Therefore,
it is not clear what $L_s$ is. While the definition of the process $L^N$ maybe clear, what is the
precise definition of $L_s$ ? I think these need to be defined through the finite-$N$ models. It may
be good to explain why the process $L$ is Markovian, or at least point out that it is Markovian,
before talking about semigroup. Probably Section \ref{sec:lineage_motion} will need to be modified accordingly.
\end{point}

\reply{
    Add sentence saying that this is well-defined because we take the finite-$N$ limit.
    Markovian because it is conditional on the population process, which is deterministic...
}

@@

\begin{point}{\revref}
% P16: Second paragraph:
The notation $L_t$ seems to first appear here and was not yet defined.
\end{point}

\reply{
    (Note: they said ``P16: Second paragraph''; I think they just missed it on the previous page.)
}

\begin{point}{}
% P19-20:
    The equations for the lineage dynamics for the three examples [i.e., Fisher-KPP \revref, Allen-Cahn, and Porous Medium] are interesting.
Which is/are new in this paper, and which was/were known before (and related references)?
\end{point}

\reply{
    The PME is new;
    probably the others are in Birzu/Korolev/Hallaschek? 
    Also maybe see Roques et al 2008ish PNAS?
    It reduces to BM in a potential:
    find other references for this (i.e., the $\grad \log$...)
}

@@

\begin{point}{}
This maybe out of the scope of this paper, but it is related. As the authors pointed out, two
lineages probably will not coalesce in the limiting model. However, they will coalesce in
the finite-$N$ model, though it may take a very long time. I wonder if the authors have a
feeling about whether the description of the lineage dynamics in the limit still be a good
approximation for that of the finite model for such a long time.
\end{point}

\reply{
    Just say we don't know, but:
    see Pennington and Etheridge on bistable waves
    where fluctuations in the wave happen on long time scales
}

@@

\reply{
}

\begin{point}{\revref}
% P65: Remark 8.1:
measure-values processes $\to$ measure-valued processes
\end{point}

\reply{
}

\begin{point}{}
Section \ref{sec:lineages_proof} may need to be modified also based on the second point in ``Major
comments'' above.
\end{point}

### 2023-12-09 forwards-and-backwards.tex
notes with alison
minor corrections from matthias

We shall write $\mu_\theta(x, X/N)$ for the per-capita death rate of mature individuals
in the population.
In order for the 
population density to change over timescales of order $\theta$, we should like the net 
per capita reproductive rate 
to scale as $1/\theta$. 
In classical models, in which $r$, $\gamma$, and $\mu$ are constant, 
this quantity is simply $r\gamma -\mu$. 
Here, because production of juveniles and their establishment are
mediated by population density measured relative to different points, 
the net reproductive rate will take a more complicated form. 
In particular, the total rate of production of mature offspring by an individual at $x$ will be
\begin{equation} \label{eqn:total_repro}
	\gamma\big(x,\smooth{\gamma}X(x)/N\big)\int r\big(y, \smooth{r}X(y)/N\big)q_{\theta}(x,dy).
\end{equation}
Nonetheless, it will be convenient to define the death rate $\mu_\theta$ in terms of 
its deviation from $r\gamma$. 
To this end, we define
the death rate of an individual at $x$ to be \revpoint{1}{4}
\begin{align} \label{eqn:mu_defn}
    \mu_\theta(x)
    =
    \max\left\{0, 
        r(x, \smooth{r} X(x) / N) \gamma(x, \smooth{\gamma} X(x) / N)
        - \frac{1}{\theta} F(x, \smooth{F} X(x) / N)
    \right\} ,
\end{align}
where this equation defines $F(x, m) : \IR^d \times \IR_{\ge 0} \to \IR$,
and $\rho_F$ is again a smooth kernel. 
The function $F$ is nearly
the net per capita reproductive rate, scaled by $\theta$,
and would be equal to it in a nonspatial model;
but, as can be seen from \eqref{eqn:total_repro},
differs because an offspring's establishment probability is measured at their new location
rather than that of their parent.
For the most part, we work with $F$ instead of $\mu_\theta$.

@@

\begin{proposition}
    \label{prop:nonlocal_to_local}
Let $\rho^\epsilon_F(x)=\rho_F\big(x/\epsilon)/\epsilon^d$.
Assume $\varphi_0\in L^2(\IR)$ is a positive, uniformly Lipschitz, and uniformly bounded function. 
Suppose that $\varphi^\epsilon\in L^2(\IR^d)$ is a weak solution to the equation
\begin{equation}
\label{nonlocalPDEv1} 
%\begin{cases}
\partial_t \varphi^\epsilon = \DG^* \varphi^\epsilon + 
\varphi^\epsilon F(\rho_F^\epsilon*\varphi^\epsilon),  
	%& 
\qquad x \in \mathbb{R}^d,\, t >0, %\\ 
%	\varphi^\epsilon(0,x) = \varphi_0(x) & x \in \mathbb{R}^d;
%\end{cases}
\end{equation}
with initial condition $\varphi_0(\cdot)$,
and that $\varphi$ is a weak solution to the equation
 \begin{equation}
\label{localPDE} 
%\begin{cases}
\partial_t \varphi = \DG^* \varphi + \varphi F(\varphi),  
%	& 
\qquad x \in \mathbb{R}^d,\, t >0, %\\ 
%	\varphi(0,x) = \varphi_0(x) & x \in \mathbb{R}^d.
%\end{cases}
\end{equation}
also with initial condition $\varphi_0(\cdot)$.
%Assume $\varphi_0\in L^2(\IR)$ is a positive, Lipschitz, and uniformly bounded function; 
Suppose further that $F$ is a Lipschitz function which is bounded above, and that
$\meanq(x)$ and $\covq(x)$, the drift and 
covariance matrix of $\DG$, satisfy the conditions of 
Assumptions~\ref{def:model_setup}.
Then, for all $T>0$ there exists a constant $K=K(T, \Vert \varphi_0 \Vert_\infty) < \infty$
and a function $\delta(\epsilon)$ (dependent on $\rho_F$) with $\delta(\epsilon)\to 0$ 
as $\epsilon\to 0$,
such that, for all $0 \leq t \leq T$, and $\epsilon$ small enough,
\[ 
\Vert \varphi_t(\cdot) - \varphi_t^\epsilon(\cdot) \Vert_\infty\leq K\delta(\epsilon). 
\]
In particular, as $\epsilon\to 0$, we have that $\varphi^\epsilon$ converges 
uniformly in compact intervals of time to $\varphi$.
\end{proposition}

@@

\begin{figure}
    \begin{center}
        % \includegraphics[width=0.7\textwidth]{figures/ex2a/pme_1234.steps}
        % \includegraphics[width=0.7\textwidth]{figures/ex2b/pme_1234.steps}
        \includegraphics[width=4.5in]{figures/fig2a}
        \includegraphics[width=4.5in]{figures/fig2b}
    \end{center}
    \caption{
        Simulated populations under a porous medium equation with logistic growth~(\ref{PME}) in $d=1$,
        % comparing the simulated profile to the analytical solution;
        $\theta/N$ small on the left; large on the right.
        % Simulation is with $\gamma(m) = m$ and $\mu(m) = (1+1/\theta) m - 1/\theta$
        % (implemented by choosing a small $dt$ taking reproduction probability per time step
        % to be $1 - \exp(- dt m)$
        % and death probability to be $\mu(m) = 1 - \exp(-dt ((1+1/\theta) m - 1/\theta))$);
        Values of $\theta$ in top and bottom figures
        are 1 and 100, respectively, and both have $N$ set
        so that the density is roughly 100 individuals per unit of habitat
        (as displayed on the vertical axis).
        See text for details of the simulations.
        \label{fig:pme_waves}
    }
\end{figure}

@@

To understand the stability of the constant solution $\varphi\equiv\varphi_0$, 
we
linearise~\eqref{general deterministic limit} around $\varphi_0$:
let $\varphi_t(x) = \varphi_0 + \psi_t(x)$,
and (informally) $r(x) \approx r(\varphi_0) + r'(\varphi_0) \smooth{r} \psi(x)$.
Writing $r_0 = r(\varphi_0)$ and $r'_0 = r'(\varphi_0)$,
with analogous expressions for $\gamma$ and $F$,
\begin{align*}
    \partial_t \psi
    &\approx
    \sigma^2 \varphi_0 r_0 \gamma_0' \Delta \smooth{\gamma} \psi
    + \sigma^2 r_0 \gamma_0 \Delta \psi
    + \varphi_0 F'_0 \smooth{F} \psi .
\end{align*}
Letting $\hat f(u) = \int e^{2\pi i u x} f(x) dx$ denote the Fourier transform,
\begin{equation}
	\label{fourier transform psi}
    \partial_t \hat \psi(u)
    \approx
	\left\{
	- u^2\sigma^2 \varphi_0 r_0 \gamma_0' \hat{\rho}_\gamma(u) 
	%e^{-\epsilon^2 u^2 /2}
        - u^2 \sigma^2 r_0 \gamma_0
	+ \varphi_0 F'_0 \hat{\rho}_F(u) 
	%e^{-\epsilon^2 u^2 /2} 
	\right\} \hat \psi(u) .
\end{equation}
In the simplest case, in which $\gamma$ is constant, so $\gamma_0'=0$,
this reduces to
\begin{equation} %\begin{align*}
	\label{simplest FT}
	\partial_t \hat \psi(u)
    \approx
    \left(
%	- u^2\sigma^2 \varphi_0 r_0 \gamma_0' \hat{\rho}_\gamma(u) 
	%e^{-\epsilon^2 u^2 /2}
        - u^2 \sigma^2 r_0 \gamma_0
	+ \varphi_0 F'_0 \hat{\rho}_F(u) 
	%e^{-\epsilon^2 u^2 /2} 
    \right) \hat \psi(u) .
\end{equation} %\end{align*}
If we take $\rho_F=p_{\epsilon^2}$, then 
$\hat{\rho}_F(u)=\exp(-2\pi\epsilon^2 u^2)$ and (recalling 
that $F_0'<0$) the term in brackets is always negative, and we recover
the well-known fact that in this case the constant solution is stable.
If, on the other hand, $\hat{\rho}_F$ changes sign, there may be values of $u$ for which the corresponding quantity is positive. For example, if $d=1$ and
$\rho_F(x)={\mathbf 1}_{[-\epsilon, \epsilon]}(x)/2\epsilon$, then 
$\hat{\rho}_F(u)=\sin (2\pi\epsilon u)/(2\pi\epsilon u)$, which is 
negative for $u\in (1/(2\epsilon), 1/\epsilon)$ (and periodically 
repeating intervals). Setting $v=\epsilon u$, the bracketed term on 
the right hand side of~\eqref{simplest FT} becomes 
\[
	\varphi_0F_0'\frac{1}{2\pi v}\sin (2\pi v)
	-\frac{\sigma^2}{\epsilon^2}v^2r_0\gamma_0,
\]
and we see that if $\sigma^2/\epsilon^2$ is sufficiently small, there are
values of $v$ for which this is positive. In other words, in keeping
with our heuristic above, if dispersal is sufficiently short range relative to
the range over which individuals interact, there are unstable frequencies that
scale with the interaction distance $\epsilon$. In two dimensions, replacing 
the indicator of an interval by that of a ball of radius $\epsilon$, a
similar analysis applies, except that the sine function is replaced by a 
Bessel function.

Now suppose that $\gamma$ is not constant. Then, 
from~\eqref{fourier transform psi},
if we take 
$\rho_\gamma=\rho_F=p_\epsilon^2$,
\begin{equation*}
	\partial_t \hat \psi(u) = e^{-2\pi^2\epsilon^2 u^2 } 
	\left\{
		-\sigma^2\varphi_0 r_0\gamma_0'u^2
	-\sigma^2r_0\gamma_0u^2e^{2\pi^2\epsilon^2u^2}
	+\varphi_0 F_0'
	\right\}
	\hat{\psi}(u).
\end{equation*}
If we make the (reasonable)
assumption that $\gamma_0'<0$, then we see that even when the Fourier
transform of $\rho$ does not change sign, there may be parameter values
for which the constant solution is unstable.
As before, we set $v=\epsilon u$. The term in brackets becomes
\[
	\frac{\sigma^2}{\epsilon^2}v^2 r_0 \left(-\varphi_0 \gamma_0' - \gamma_0 e^{2\pi^2v^2}\right)+\varphi_0F_0',
\] 
and, provided $-\varphi_0\gamma_0'/\gamma_0>1$, for sufficiently small $v$ 
the term in round brackets is positive. We now see that if 
$\sigma^2/\epsilon^2$ is sufficiently {\em large},
the equilibrium state $\varphi\equiv\varphi_0$ 
is unstable. As before, the unstable frequencies will scale with $\epsilon$
and for given $F$, $r$ and $\gamma$, whether or not such unstable
frequencies exist will be determined by $\sigma^2/\epsilon^2$, but 
in this case of Gaussian kernels, 
it is interaction distance being sufficiently small relative
to dispersal that will lead to instability.

@@

Next we must choose the new level created at the birth event.
We would like an individual with level $u$ and at spatial position $x$
to produce offspring at $y$ at instantaneous rate
\begin{equation}
	\label{rate of birth}
	2 \Big(1 - \frac{u}{N}\Big) \theta \gamma(x, \eta) r(x + y, \eta).
\end{equation}
To do this
we will associate the point $(\tau, v, z, \kappa) \in \Pi_a$ with 
level $u + v \ell$,
where $\ell$ is chosen so that the rate of appearance of points in 
$\Pi_a$ with level below $N$, that is points with
$v \ell <N-u$, is given by~(\ref{rate of birth}).
Since the mean measure of $\Pi_a$ is Lebesgue measure
in the $t$ and $v$ directions, we must take
\begin{equation}
	\label{expression for ell}
    \ell(x, y, \eta)
    =
    \frac{N-u}{2 (1-u/N) \theta \gamma(x, \eta) r(x + y, \eta) } 
    =\frac{1}{2N^{-1} \theta \gamma(x, \eta) r(x + y, \eta) } 
    ,
\end{equation}
and, using this, the (potential) new level is
\begin{equation*}
    U_{a \concat j}(\tau)
    =
    U_a(\tau)
    +
    v \ell\big(X_a(\tau-), y(X_a(\tau-), z), \eta_{\tau-} \big) .
\end{equation*}
If $U_{a \concat j}(\tau) < N$,
the new individual labeled $a \concat j$ is produced,
and $\kappa$ determines which label, $a$ or $a\concat j$,
is associated with the new location,
so
\begin{align*}
    X_{a \concat j}(\tau)
    &=
    X_a(\tau-) + (1 - \kappa) y\big(X_a(\tau-), z\big) .
\end{align*}
On the other hand
if $U_{a \concat j}(\tau) \ge N$, then $X_a$ is unchanged and 
$X_{a \concat j}$ is undefined,
so
\begin{align} \label{eqn:X_line_of_descent}
    X_a(\tau)
    &=
    X_a(\tau-) + \kappa y(X_a(\tau-), z) \ind_{U_{a \concat j}(\tau) < N} .
\end{align}
Recall that the parental \emph{individual} always retains their spatial location,
so that $\kappa = 0$ corresponds to the parent being assigned a new level,
and our line of descent switching to the offspring.
Combining these observations, $X_a$, for $\tau_a \le t < \sigma_a$, solves the equation 
\begin{align*}
    X_a(t)
    &=
    X_a(\tau_a) + 
    \int_{[\tau_a, t) \times [0, \infty) \times \IR \times [0, 1]}
    y(X_a(\tau-), z)
    \kappa
    \ind_{
        U_a(\tau) + v \ell(X_a(\tau-), y(X_a(\tau-), z), \eta_{\tau-})<N
    }
    d\Pi_a(\tau, v, z, \kappa) .
\end{align*}

@@

We rewrite $\Pgen^NG_f(\eta^N_s)$ using a 
Taylor series expansion up to third order for
$G\big(\langle f,\eta\rangle\pm f(y)/N\big)$ around
$G(\langle f, \eta \rangle )$.
As in Section~\ref{sec:heuristics} (except that now we are more explicit about
the error term), we find
\begin{equation} 
    \label{eq: Pre-Limit Generator Expanded}
\begin{aligned}
\Pgen^{N} G_f(\eta)
    :=& 
    G'(\langle f, \eta \rangle)\int_{\IR^d}  
\theta\Big\langle \gamma(x,\eta) \int_{\IR^d} f(y)r(y,\eta) q_{\theta}(x,dy), 
    - f(x)\mu_\theta(x,\eta), \eta(dx)\Big\rangle\\
&+\frac{1}{2}\frac{\theta}{N}G''(\langle f, \eta \rangle)\int_{\IR^d} 
 \Big\langle\gamma(x, \eta)\int_{\IR^d} f^2(y) r(y,\eta)q_{\theta}(x,dy)+f^2(x)
\mu_\theta(x, \eta), \eta(dx)\Big\rangle\\
&+\frac{1}{6}\frac{\theta}{N^2}\Big\langle
G'''(w) \gamma(x, \eta)\int_{\IR^d} f^3(y) r(y,\eta) 
q_{\theta}(x,dy)
-G'''(v)f^3(x)\mu_\theta(x, \eta), \eta(dx)\Big\rangle
\end{aligned}    
\end{equation}
for some $w,v \in [\langle f,\eta \rangle - \|f\|_\infty/N, 
\langle f,\eta \rangle + \|f\|_\infty/N]$.

@@

It is natural to conjecture that when the limit of the rescaled 
population process that we obtained in the previous section solves a nonlocal
PDE, if we further scale the kernels 
$\rho_r$, $\rho_\gamma$, and $\rho_F$ by setting 
$\rho^\epsilon(\cdot)=\rho(\cdot/\epsilon)/\epsilon$, as $\epsilon\to 0$, the 
corresponding solutions should converge to a limiting
population density that solves the corresponding ``classical'' PDE. We verify
this in two examples; in the first the nonlocal equation is a reaction-diffusion equation
with the ``nonlocality'' only appearing in the 
reaction term; in the second the nonlocal PDE is a special 
case of a nonlinear porous medium equation. These, in particular,
capture the examples that we explored in 
Section~\ref{ancestral lineages for travelling waves}.

@@

\begin{remark}
In addition to tightness of the measure-values processes $\lp^N$, \revpoint{1}{34}
the bounds used in the proofs below also imply
tightness of the number of lines of descent and the number of births below a fixed level,
and of the motion of individual lines of descent.
In other words, the limiting ``line of descent'' construction
of Section~\ref{sec: individual lines of descent} holds.
\end{remark}

@@

% \bibliographystyle{amsplain}
% \bibliography{refs}

@@

\providecommand{\bysame}{\leavevmode\hbox to3em{\hrulefill}\thinspace}
\providecommand{\MR}{\relax\ifhmode\unskip\space\fi MR }
% \MRhref is called by the amsart/book/proc definition of \MR.
\providecommand{\MRhref}[2]{%
  \href{http://www.ams.org/mathscinet-getitem?mr=#1}{#2}
}
\providecommand{\href}[2]{#2}
\begin{thebibliography}{10}

>>>>

We shall write $\mu_\theta(x, X/N)$ for the per-capita death rate of a mature individual
at $x$ in the population.
In order for the 
population density to change over timescales of order $\theta$, we should like the net 
per capita reproductive rate 
to scale as $1/\theta$. 
In classical models, in which $r$, $\gamma$, and $\mu$ are constant, 
this quantity is simply $r\gamma -\mu$. 
Here, because production of juveniles and their establishment are
mediated by population density measured relative to different points, 
the net reproductive rate will take a more complicated form. 
In particular, the total rate of production of mature offspring by an individual at $x$ will be
\begin{equation} \label{eqn:total_repro}
	\gamma\big(x,\smooth{\gamma}X(x)/N\big)\int r\big(y, \smooth{r}X(y)/N\big)q_{\theta}(x,dy).
\end{equation}
Nonetheless, it will be convenient to define the death rate $\mu_\theta$ in terms of 
its deviation from $r\gamma$. 
To this end, we define
the death rate of an individual at $x$ to be \revpoint{1}{4}
\begin{align} \label{eqn:mu_defn}
    \mu_\theta(x, X/N)
    =
    \max\left\{0, 
        r(x, \smooth{r} X(x) / N) \gamma(x, \smooth{\gamma} X(x) / N)
        - \frac{1}{\theta} F(x, \smooth{F} X(x) / N)
    \right\} ,
\end{align}
where this equation also introduces $F(x, m) : \IR^d \times \IR_{\ge 0} \to \IR$,
and $\rho_F$ is again a smooth kernel. 
The function $F$ is nearly
the net per capita reproductive rate, scaled by $\theta$,
and would be equal to it in a nonspatial model;
but, as can be seen from \eqref{eqn:total_repro},
differs because an offspring's establishment probability is measured at their new location
rather than that of their parent.
For the most part, we work with $F$ instead of $\mu_\theta$.

@@

\begin{proposition}
    \label{prop:nonlocal_to_local}
Let $\rho^\epsilon_F(x)=\rho_F\big(x/\epsilon)/\epsilon^d$.
Assume $\varphi_0\in L^2(\IR^d)$ is a positive, uniformly Lipschitz, and uniformly bounded function. 
Suppose that $\varphi^\epsilon\in L^2(\IR^d)$ is a weak solution to the equation
\begin{equation}
\label{nonlocalPDEv1} 
%\begin{cases}
\partial_t \varphi^\epsilon = \DG^* \varphi^\epsilon + 
\varphi^\epsilon F(\rho_F^\epsilon*\varphi^\epsilon),  
	%& 
\qquad x \in \mathbb{R}^d,\, t >0, %\\ 
%	\varphi^\epsilon(0,x) = \varphi_0(x) & x \in \mathbb{R}^d;
%\end{cases}
\end{equation}
with initial condition $\varphi_0(\cdot)$,
and that $\varphi$ is a weak solution to the equation
 \begin{equation}
\label{localPDE} 
%\begin{cases}
\partial_t \varphi = \DG^* \varphi + \varphi F(\varphi),  
%	& 
\qquad x \in \mathbb{R}^d,\, t >0, %\\ 
%	\varphi(0,x) = \varphi_0(x) & x \in \mathbb{R}^d.
%\end{cases}
\end{equation}
also with initial condition $\varphi_0(\cdot)$.
%Assume $\varphi_0\in L^2(\IR)$ is a positive, Lipschitz, and uniformly bounded function; 
Suppose further that $F$ is a Lipschitz function which is bounded above, and that
$\meanq(x)$ and $\covq(x)$, the drift and 
covariance matrix of $\DG$, satisfy the conditions of 
Assumptions~\ref{def:model_setup}.
Then, for all $T>0$ there exists a constant $K=K(T, \Vert \varphi_0 \Vert_\infty) < \infty$
and a function $\delta(\epsilon)$ (dependent on $\rho_F$) with $\delta(\epsilon)\to 0$ 
as $\epsilon\to 0$,
such that, for all $0 \leq t \leq T$, and $\epsilon$ small enough,
\[ 
\Vert \varphi_t(\cdot) - \varphi_t^\epsilon(\cdot) \Vert_\infty\leq K\delta(\epsilon). 
\]
In particular, as $\epsilon\to 0$, we have that $\varphi^\epsilon$ converges 
uniformly in compact intervals of time to $\varphi$.
\end{proposition}

@@

\begin{figure}
    \begin{center}
        % \includegraphics[width=0.7\textwidth]{figures/ex2a/pme_1234.steps}
        % \includegraphics[width=0.7\textwidth]{figures/ex2b/pme_1234.steps}
        \includegraphics[width=4.5in]{figures/fig2a}
        \includegraphics[width=4.5in]{figures/fig2b}
    \end{center}
    \caption{
        Simulated populations under a porous medium equation with logistic growth~(\ref{PME}) in $d=1$,
        % comparing the simulated profile to the analytical solution;
        $\theta/N$ small on the top; large on the bottom.
        % Simulation is with $\gamma(m) = m$ and $\mu(m) = (1+1/\theta) m - 1/\theta$
        % (implemented by choosing a small $dt$ taking reproduction probability per time step
        % to be $1 - \exp(- dt m)$
        % and death probability to be $\mu(m) = 1 - \exp(-dt ((1+1/\theta) m - 1/\theta))$);
        Values of $\theta$ in top and bottom figures
        are 1 and 100, respectively, and both have $N$ set
        so that the density is roughly 100 individuals per unit of habitat
        (as displayed on the vertical axis).
        See text for details of the simulations.
        \label{fig:pme_waves}
    }
\end{figure}

@@

To understand the stability of the constant solution $\varphi\equiv\varphi_0$, 
we
linearise~\eqref{general deterministic limit} around $\varphi_0$:
let $\varphi_t(x) = \varphi_0 + \psi_t(x)$,
and (informally) $r(x) \approx r(\varphi_0) + r'(\varphi_0) \smooth{r} \psi(x)$.
Writing $r_0 = r(\varphi_0)$ and $r'_0 = r'(\varphi_0)$,
with analogous expressions for $\gamma$ and $F$,
\begin{align*}
    \partial_t \psi
    &\approx
    \sigma^2 \varphi_0 r_0 \gamma_0' \Delta \smooth{\gamma} \psi
    + \sigma^2 r_0 \gamma_0 \Delta \psi
    + \varphi_0 F'_0 \smooth{F} \psi .
\end{align*}
Letting $\hat f(u) = \int e^{i u x} f(x) dx / \sqrt{2 \pi}$ denote the Fourier transform,
\begin{equation}
	\label{fourier transform psi}
    \partial_t \hat \psi(u)
    \approx
	\left\{
	- u^2\sigma^2 \varphi_0 r_0 \gamma_0' \hat{\rho}_\gamma(u) 
	%e^{-\epsilon^2 u^2 /2}
        - u^2 \sigma^2 r_0 \gamma_0
	+ \varphi_0 F'_0 \hat{\rho}_F(u) 
	%e^{-\epsilon^2 u^2 /2} 
	\right\} \hat \psi(u) .
\end{equation}
In the simplest case, in which $\gamma$ is constant, so $\gamma_0'=0$,
this reduces to
\begin{equation} %\begin{align*}
	\label{simplest FT}
	\partial_t \hat \psi(u)
    \approx
    \left(
%	- u^2\sigma^2 \varphi_0 r_0 \gamma_0' \hat{\rho}_\gamma(u) 
	%e^{-\epsilon^2 u^2 /2}
        - u^2 \sigma^2 r_0 \gamma_0
	+ \varphi_0 F'_0 \hat{\rho}_F(u) 
	%e^{-\epsilon^2 u^2 /2} 
    \right) \hat \psi(u) .
\end{equation} %\end{align*}
If we take $\rho_F=p_{\epsilon^2}$, then 
$\hat{\rho}_F(u)=\exp(-\epsilon^2 u^2 / 2) / \sqrt{2 \pi}$ and (recalling 
that $F_0'<0$) the term in brackets is always negative, and we recover
the well-known fact that in this case the constant solution is stable.
If, on the other hand, $\hat{\rho}_F$ changes sign, there may be values of $u$ for which the corresponding quantity is positive. For example, if $d=1$ and
$\rho_F(x)={\mathbf 1}_{[-\epsilon, \epsilon]}(x)/2\epsilon$, then 
$\hat{\rho}_F(u)=\sin(\epsilon u)/(\sqrt{2\pi}\epsilon u)$, which is 
negative for $u\in (\pi/epsilon, 2\pi/\epsilon)$ (and periodically 
repeating intervals). Setting $v=\epsilon u$, the bracketed term on 
the right hand side of~\eqref{simplest FT} becomes 
\[
    \varphi_0F_0'\frac{1}{\sqrt{2\pi} v} \sin(v)
	-\frac{\sigma^2}{\epsilon^2}v^2r_0\gamma_0,
\]
and we see that if $\sigma^2/\epsilon^2$ is sufficiently small, there are
values of $v$ for which this is positive. In other words, in keeping
with our heuristic above, if dispersal is sufficiently short range relative to
the range over which individuals interact, there are unstable frequencies that
scale with the interaction distance $\epsilon$. In two dimensions, replacing 
the indicator of an interval by that of a ball of radius $\epsilon$, a
similar analysis applies, except that the sine function is replaced by a 
Bessel function.

Now suppose that $\gamma$ is not constant. Then, 
from~\eqref{fourier transform psi},
if we take 
$\rho_\gamma=\rho_F=p_\epsilon^2$,
\begin{equation*}
	\partial_t \hat \psi(u) =
    \frac{1}{\sqrt{2\pi}}
    e^{-\epsilon^2 u^2/2} 
	\left\{
		-\sigma^2 \varphi_0 r_0 \gamma_0' u^2
        -\sigma^2 r_0 \gamma_0 u^2 \sqrt{2\pi} e^{\epsilon^2 u^2/2}
        +\varphi_0 F_0'
	\right\}
	\hat{\psi}(u).
\end{equation*}
If we make the (reasonable)
assumption that $\gamma_0'<0$, then we see that even when the Fourier
transform of $\rho$ does not change sign, there may be parameter values
for which the constant solution is unstable.
As before, we set $v=\epsilon u$. The term in brackets becomes
\[
	\frac{\sigma^2}{\epsilon^2}v^2 r_0
    \left(
        -\varphi_0 \gamma_0' - \gamma_0 \sqrt{2\pi} e^{v^2/2}
    \right) + \varphi_0F_0',
\] 
and, provided $-\varphi_0\gamma_0'/\gamma_0\sqrt{2\pi} >1$, for sufficiently small $v$ 
the term in round brackets is positive. We now see that if 
$\sigma^2/\epsilon^2$ is sufficiently {\em large},
the equilibrium state $\varphi\equiv\varphi_0$ 
is unstable. As before, the unstable frequencies will scale with $\epsilon$
and for given $F$, $r$ and $\gamma$, whether or not such unstable
frequencies exist will be determined by $\sigma^2/\epsilon^2$, but 
in this case of Gaussian kernels, 
it is interaction distance being sufficiently small relative
to dispersal that will lead to instability.

@@

Next we must choose the new level created at the birth event.
We would like an individual with level $u$ and at spatial position $x$
to produce offspring at $y$ at instantaneous rate
\begin{equation}
	\label{rate of birth}
	2 \Big(1 - \frac{u}{N}\Big) \theta \gamma(x, \eta) r(x + y, \eta).
\end{equation}
To do this
we will associate the point $(\tau, v, z, \kappa) \in \Pi_a$ with 
level $u + v \ell$,
where $\ell$ is chosen so that the rate of appearance of points in 
$\Pi_a$ with level below $N$, that is points with
$v \ell <N-u$, is given by~(\ref{rate of birth}).
Since the mean measure of $\Pi_a$ is Lebesgue measure
in the $t$ and $v$ directions, we must take
\begin{equation}
	\label{expression for ell}
    \ell(x, y, \eta)
    =
    \frac{N-u}{2 (1-u/N) \theta \gamma(x, \eta) r(x + y, \eta) } 
    =\frac{1}{2N^{-1} \theta \gamma(x, \eta) r(x + y, \eta) } 
    ,
\end{equation}
and, using this, the (potential) new level is
\begin{equation*}
    U_{a \concat j}(\tau)
    =
    U_a(\tau)
    +
    v \ell\big(X_a(\tau-), y(X_a(\tau-), z), \eta_{\tau-} \big) .
\end{equation*}
If $U_{a \concat j}(\tau) < N$,
the new individual labeled $a \concat j$ is produced,
and $\kappa$ determines which label, $a$ or $a\concat j$,
is associated with the new location,
so
\begin{align*}
    X_{a \concat j}(\tau)
    &=
    X_a(\tau-) + (1 - \kappa) y\big(X_a(\tau-), z\big) .
\end{align*}
On the other hand
if $U_{a \concat j}(\tau) \ge N$, then $X_a$ is unchanged and 
$X_{a \concat j}$ is undefined,
so
\begin{align} \label{eqn:X_line_of_descent}
    X_a(\tau)
    &=
    X_a(\tau-) + \kappa y(X_a(\tau-), z) \ind_{U_{a \concat j}(\tau) < N} .
\end{align}
Recall that the parental \emph{individual} always retains their spatial location,
so that $\kappa = 1$ corresponds to the parent being assigned a new level,
and our line of descent switching to the offspring.
Combining these observations, $X_a$, for $\tau_a \le t < \sigma_a$, solves the equation 
\begin{align*}
    X_a(t)
    &=
    X_a(\tau_a) + 
    \int_{[\tau_a, t) \times [0, \infty) \times \IR \times [0, 1]}
    y(X_a(\tau-), z)
    \kappa
    \ind_{
        U_a(\tau) + v \ell(X_a(\tau-), y(X_a(\tau-), z), \eta_{\tau-})<N
    }
    d\Pi_a(\tau, v, z, \kappa) .
\end{align*}

@@

We rewrite $\Pgen^NG_f(\eta^N_s)$ using a 
Taylor series expansion up to third order for
$G\big(\langle f,\eta\rangle\pm f(y)/N\big)$ around
$G(\langle f, \eta \rangle )$.
As in Section~\ref{sec:heuristics} (except that now we are more explicit about
the error term), we find
\begin{equation} 
    \label{eq: Pre-Limit Generator Expanded}
\begin{aligned}
\Pgen^{N} G_f(\eta)
    :=& 
    G'(\langle f, \eta \rangle)\int_{\IR^d}  
\theta\Big\langle \gamma(x,\eta) \int_{\IR^d} f(y)r(y,\eta) q_{\theta}(x,dy) 
    - f(x)\mu_\theta(x,\eta), \eta(dx)\Big\rangle\\
&+\frac{1}{2}\frac{\theta}{N}G''(\langle f, \eta \rangle)\int_{\IR^d} 
 \Big\langle\gamma(x, \eta)\int_{\IR^d} f^2(y) r(y,\eta)q_{\theta}(x,dy)+f^2(x)
\mu_\theta(x, \eta), \eta(dx)\Big\rangle\\
&+\frac{1}{6}\frac{\theta}{N^2}\Big\langle
G'''(w) \gamma(x, \eta)\int_{\IR^d} f^3(y) r(y,\eta) 
q_{\theta}(x,dy)
-G'''(v)f^3(x)\mu_\theta(x, \eta), \eta(dx)\Big\rangle
\end{aligned}    
\end{equation}
for some $w,v \in [\langle f,\eta \rangle - \|f\|_\infty/N, 
\langle f,\eta \rangle + \|f\|_\infty/N]$.

@@

It is natural to conjecture that when the limit of the rescaled 
population process that we obtained in the previous section solves a nonlocal
PDE, if we further scale the kernels 
$\rho_r$, $\rho_\gamma$, and $\rho_F$ by setting 
$\rho^\epsilon(\cdot)=\rho(\cdot/\epsilon)/\epsilon^d$, as $\epsilon\to 0$, the 
corresponding solutions should converge to a limiting
population density that solves the corresponding ``classical'' PDE. We verify
this in two examples; in the first the nonlocal equation is a reaction-diffusion equation
with the ``nonlocality'' only appearing in the 
reaction term; in the second the nonlocal PDE is a special 
case of a nonlinear porous medium equation. These, in particular,
capture the examples that we explored in 
Section~\ref{ancestral lineages for travelling waves}.

@@

\begin{remark}
In addition to tightness of the measure-valued processes $\lp^N$, \revpoint{1}{34}
the bounds used in the proofs below also imply
tightness of the number of lines of descent and the number of births below a fixed level,
and of the motion of individual lines of descent.
In other words, the limiting ``line of descent'' construction
of Section~\ref{sec: individual lines of descent} holds.
\end{remark}

@@

\bibliographystyle{amsplain}
\bibliography{refs}

@@

% \providecommand{\bysame}{\leavevmode\hbox to3em{\hrulefill}\thinspace}
% \providecommand{\MR}{\relax\ifhmode\unskip\space\fi MR }
% % \MRhref is called by the amsart/book/proc definition of \MR.
% \providecommand{\MRhref}[2]{%
%   \href{http://www.ams.org/mathscinet-getitem?mr=#1}{#2}
% }
% \providecommand{\href}[2]{#2}
% \begin{thebibliography}{10}
% 
